\documentclass[12pt,a4paper]{article}

\usepackage{amsmath}
\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\setlength\topmargin{0in}
\setlength\headheight{0in}
\setlength\headsep{0in}
\setlength\textheight{10in}
\setlength\textwidth{6.5in}
\setlength\oddsidemargin{0in}
\setlength\evensidemargin{0in}

\title{Data Mining project status report}
\author{Gasparella Luca, Sartori Enrico}
\date{A.y. 2008/2009}

\begin{document}
\maketitle
\subsection*{Details on problem definition}
The main aim of our project is to provide an implementation of the Random
Decision Tree approach proposed in 2003 by Wei Fan \ref{rdt}, and compare
its results and performance with a normal classifier.

The main dataset which we are using to develop the project is taken from
the 2008 ICDM Data Mining contest, and represents the
concentration of a set of Xenon isotopes samples. These samples are used
to find out possible unauthorized nuclear tests.

From the implementation point of view we choose as main language Python,
due to its great readability and because it give us the possibility to
focus more on the algorithm developing part than on more technical issues.
Anyway, we consider the possibility to integrate some C sections in our
software, for dealing with the most performance critical parts.

\subsection*{Plan Revision}
As revision of the project proposal, we consider as aim of the project not
only provide an implementation of the RDT approach, but perform a complete
set of test in order to prove and evaluate the reliability of this
approach in comparison with a normal classifier.

For this reason we should consider more than one single dataset, in order
to evaluate the performances of the different approaches in various
situation.

In order to have the possibility to operate on different datasets we
choose to adopt a common format of data representation, we selected the
Arff format. We built some tools useful to provide support for converting
the datasets on which we are working in arff files.

For optimize the I/O operations, we decided to read the date from disk in
blocks which can fit in main memory easily. In this way we don't need to
access to the disk for every datapoint, but they are buffered in main
memory.   

\subsection*{Evaluation approach}
For our project we have three main metrics to check in order to evaluate
the performances of the algorithms: accuracy, time and memory consumption.
For accuracy we intend the quality of the classification obtained, which
is measurable using cross validation techniques, or, in the case of the
Xenon concentration dataset, using a specific application provided by the
contest organizers.

This application checks the classification obtained for the test dataset,
giving a measure of the quality obtained.

\subsection*{Status}
We have analyzed the contests dataset and found some differences of the
scale used by the different stations to represent the gas concentration,
so we added an attribute useful to identify the station and adapt the
classifier to the scale used.

As we use new datasets we will provide some tools for converting them into
arff files, if needed. At the moment we developed a tool for converting
the contests datasets into arff files.

In this moment we are focusing on the implementation of a little framework
which can manage the execution of different algorithms, in order to ease
the performance evaluation task.

\begin{thebibliography}{9}
\bibitem{rdt} Wei Fan, Haixun Wang, Philip S. Yu, Sheng Ma  \emph{"Is
random model better? On its accuracy and efficiency"},  3rd IEEE
International Conference on Data Mining (ICDM 2003), November 19-22,
Melbourne, FL: 51-58. 

\bibitem{opt} Wei Fan  \emph{"On the Optimality of Probability Estimation
by Random Decision Trees"},  The Nineteenth National Conference on
Artificial Intelligence (AAAI'04), San Jose, CA, July, 2004.

\bibitem{book} \emph{ICDM Data Mining Contest '08 booklet},\\
 http://www.cs.uu.nl/groups/ADA/icdm08cup/booklet.pdf.

\end{thebibliography}

\end{document}

